Online Multiclass Linear Classification with Bandit Feedback

### 2018-12-19 arxiv-version/main.tex
Finish margin separation for rational function

\begin{proposition}[Properties of $P_n$]
\label{proposition:properties-of-p-n}
\hspace{1cm} % Dummy space
\begin{enumerate}
\item If $z \in [0,1]$ then $P_n(-z) \le P_n(z) \le 0$.
\item If $z \in [1,2^n]$ then $0 \le 4P_n(z) \le -P_n(-z)$.
\item If $z \ge 0$ then $-P_n(-z) \ge 2^{n(n+1)}$.
\end{enumerate}
\end{proposition}

@@

\begin{proposition}[Properties of $S_{n,r}$ and $B_{n,r}$]
\label{proposition:properties-of-s-n-r}
For any integer $n \ge 0$ and $r = 2 \left\lceil \frac{1}{4} \log_2(4m + 1) \right\rceil + 1$,
\begin{enumerate}
\item If $z \in [1,2^n]$ then $S_{n,r}(z) \in [1,1+\frac{1}{2m}]$.
\item If $z \in [-2^n, -1]$ then $S_{n,r} \in [-1-\frac{1}{2m}, -1]$.
\item If $z \in [-1,1]$ then $|S_{n,r}(z)| \le 1$.
\item If $z \in [-2^n,2^n]$ then $B_{n,r}(z) \ge \left(1 - \frac{1}{4m+1} \right) 2^{n(n+1)r}$.
\end{enumerate}
\end{proposition}

@@

\begin{enumerate}
\item Observe that $S_{n,r}(z)$ can be written as
$$
S_{n,r}(z) = \frac{\displaystyle 1 + \left( - \frac{P_n(z)}{P_n(-z)}\right)^r}{\displaystyle 1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r} = \frac{1 + c}{1 - c}
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. Since $z \in [1,2^n]$, by
part 2 of~\autoref{proposition:properties-of-p-n}, $c \in [0,\frac{1}{4^r}]$. Since
$r \ge \frac{1}{2} \log_2(4m+1)$ that means that $c \in [0,\frac{1}{4m+1}]$. Thus,
$S_{n,r}(z) = \frac{1+c}{1-c} \in [1,1 + \frac{1}{2m}]$.

@@

\item Observe that $S_{n,r}(z)$ can be written as
$$
S_{n,r}(z) = \frac{\displaystyle 1 + \left( - \frac{P_n(z)}{P_n(-z)}\right)^r}{\displaystyle 1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r} = \frac{1 + c}{1 - c}
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. If $z \in [0,1]$, by part
1 of~\autoref{proposition:properties-of-p-n} and the fact that $r$ is odd, $c
\in [-1,0]$, and thus, $S_{n,r}(z) = \frac{1+c}{1-c} \in [0,1]$. Since
$S_{n,r}(z)$ is odd, for $z \in [-1,0]$, $S_{n,r}(z) \in [-1,0]$.

@@

Case $z \in [0,1]$. Since $r$ is odd and $P_n(z)$ is non-positive,
$$
B_{n,r}(z) = - (P_n(z))^r + \left(- P_{n}(-z)\right)^r \ge \left(- P_{n}(-z)\right)^r \ge 2^{n(n+1)r} \ge 2^{n(n+1)r} \left( 1 - \frac{1}{4m+1} \right) \; .
$$
where the second last inequality follows from part 3 of \autoref{proposition:properties-of-p-n}.

Case $z \in [1,2^n]$. Since $r$ is odd,
$$
B_{n,r}(z) = \left(- P_{n}(-z)\right)^r \left(1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r \right) = \left(- P_{n}(-z)\right)^r (1 - c)
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. Since $z \in [1,2^n]$, by
part 2 of~\autoref{proposition:properties-of-p-n}, $c \in [0,\frac{1}{4^r}]$. By
the definition of $r$ that means that $c \in [0,\frac{1}{4m+1}]$. Thus,
$$
B_{n,r}(z) \ge \left(- P_{n}(-z)\right)^r \left( 1 - \frac{1}{4m+1} \right) \ge 2^{n(n+1)r} \left( 1 - \frac{1}{4m+1} \right) \; .
$$
where the last inequality follows from part 3 of \autoref{proposition:properties-of-p-n}.
\end{enumerate}
\end{proof}

To prove parts 1 and 2 of \autoref{theorem:polynomial-approximation-1}
note that \autoref{proposition:properties-of-s-n-r} implies that
\begin{align*}
|p(x)|
& = |Q(x)| \prod_{j=1}^m B_{s,r} \left( \frac{\ip{v_j}{x}}{\gamma} \right) \\
& \ge |Q(x)| \left( 2^{s(s+1)r} \left( 1 - \frac{1}{4m+1} \right) \right)^m \\
& \ge |Q(x)| \cdot 2^{s(s+1)rm} e^{-4} && \text{(since $\left(1-\frac{1}{4m+1}\right)^m \ge e^{-4}$)} \\
& \ge 4 |Q(x)| \cdot 2^{s^2rm} && \text{(since $e^{-4} \ge 1/2$ and $2^{rms} \ge 8$)} \\
& \ge 4 |Q(x)| \cdot (4m+1)^{s^2m/2} && \text{(since $r \ge 1 + \frac{1}{2} \log_2(4m+1)$)} \\
& \ge 2 |Q(x)| \cdot m^{s^2m/2} \; .
\end{align*}
It thus remains to prove that
\begin{align}
Q(x) & \ge \frac{1}{2} && \text{for all} \ x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\} \; , \\
Q(x) & \le -\frac{1}{2} && \text{for all} \ x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\} \; .
\end{align}

>>>>

\begin{lemma}[Properties of $P_n$]
\label{lemma:properties-of-p-n}
\hspace{1cm} % Dummy space
\begin{enumerate}
\item If $z \in [0,1]$ then $P_n(-z) \le P_n(z) \le 0$.
\item If $z \in [1,2^n]$ then $0 \le 4P_n(z) \le -P_n(-z)$.
\item If $z \ge 0$ then $-P_n(-z) \ge 2^{n(n+1)}$.
\end{enumerate}
\end{lemma}

@@

\begin{lemma}[Properties of $S_{n,r}$ and $B_{n,r}$]
\label{lemma:properties-of-s-n-r}
Let $n,m$ be non-negative integers.
Let $r = 2 \left\lceil \frac{1}{4} \log_2(4m + 1) \right\rceil + 1$. Then,
\begin{enumerate}
\item If $z \in [1,2^n]$ then $S_{n,r}(z) \in [1,1+\frac{1}{2m}]$.
\item If $z \in [-2^n, -1]$ then $S_{n,r} \in [-1-\frac{1}{2m}, -1]$.
\item If $z \in [-1,1]$ then $|S_{n,r}(z)| \le 1$.
\item If $z \in [-2^n,2^n]$ then $B_{n,r}(z) \ge \left(1 - \frac{1}{4m+1} \right) 2^{n(n+1)r}$.
\end{enumerate}
\end{lemma}

@@

\begin{enumerate}
\item Observe that $S_{n,r}(z)$ can be written as
$$
S_{n,r}(z) = \frac{\displaystyle 1 + \left( - \frac{P_n(z)}{P_n(-z)}\right)^r}{\displaystyle 1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r} = \frac{1 + c}{1 - c}
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. Since $z \in [1,2^n]$, by
part 2 of~\autoref{lemma:properties-of-p-n}, $c \in [0,\frac{1}{4^r}]$. Since
$r \ge \frac{1}{2} \log_2(4m+1)$ that means that $c \in [0,\frac{1}{4m+1}]$. Thus,
$S_{n,r}(z) = \frac{1+c}{1-c} \in [1,1 + \frac{1}{2m}]$.

@@

\item Observe that $S_{n,r}(z)$ can be written as
$$
S_{n,r}(z) = \frac{\displaystyle 1 + \left( - \frac{P_n(z)}{P_n(-z)}\right)^r}{\displaystyle 1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r} = \frac{1 + c}{1 - c}
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. If $z \in [0,1]$, by part
1 of~\autoref{lemma:properties-of-p-n} and the fact that $r$ is odd, $c
\in [-1,0]$, and thus, $S_{n,r}(z) = \frac{1+c}{1-c} \in [0,1]$. Since
$S_{n,r}(z)$ is odd, for $z \in [-1,0]$, $S_{n,r}(z) \in [-1,0]$.

@@

Case $z \in [0,1]$. Since $r$ is odd and $P_n(z)$ is non-positive,
$$
B_{n,r}(z) = - (P_n(z))^r + \left(- P_{n}(-z)\right)^r \ge \left(- P_{n}(-z)\right)^r \ge 2^{n(n+1)r} \ge 2^{n(n+1)r} \left( 1 - \frac{1}{4m+1} \right) \; .
$$
where the second last inequality follows from part 3 of \autoref{lemma:properties-of-p-n}.

Case $z \in [1,2^n]$. Since $r$ is odd,
$$
B_{n,r}(z) = \left(- P_{n}(-z)\right)^r \left(1 - \left( - \frac{P_n(z)}{P_n(-z)}\right)^r \right) = \left(- P_{n}(-z)\right)^r (1 - c)
$$
where $c = \left( - \frac{P_n(z)}{P_n(-z)}\right)^r$. Since $z \in [1,2^n]$, by
part 2 of~\autoref{lemma:properties-of-p-n}, $c \in [0,\frac{1}{4^r}]$. By
the definition of $r$ that means that $c \in [0,\frac{1}{4m+1}]$. Thus,
$$
B_{n,r}(z) \ge \left(- P_{n}(-z)\right)^r \left( 1 - \frac{1}{4m+1} \right) \ge 2^{n(n+1)r} \left( 1 - \frac{1}{4m+1} \right) \; .
$$
where the last inequality follows from part 3 of \autoref{lemma:properties-of-p-n}.
\end{enumerate}
\end{proof}

\begin{lemma}[Properties of $Q(x)$]
\label{lemma:properties-of-q}
The rational function $Q(x)$ satisfies
\begin{enumerate}
\item $Q(x) \ge \frac{1}{2}$ for all $\displaystyle x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$,
\item $Q(x) \le -\frac{1}{2}$ for all $\displaystyle x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$.
\end{enumerate}
\end{lemma}

\begin{proof}
To prove part 1, consider any $x \in \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \
\ip{v_i}{x} \ge \gamma \right\}$. Then, $\frac{\ip{v_i}{x}}{\gamma} \in [1,
\frac{1}{\gamma}]$. By part 1 of \autoref{lemma:properties-of-s-n-r},
$S_{s,r}\left(\frac{\ip{v_i}{x}}{\gamma}\right) \in [1, 1 + \frac{1}{2m}]$ and
in particular $S_{s,r}\left(\frac{\ip{v_i}{x}}{\gamma}\right) \ge 1$. Thus,
$$
Q(x) = \left( \sum_{i=1}^m S_{s,r}\left(\frac{\ip{v_i}{x}}{\gamma}\right) \right) - (m - 1/2) \ge m - (m - 1/2) = 1/2 \; .
$$

To prove part 2, consider any $x \in \bigcup_{i=1}^m \left\{ x \in \R^d ~:~
\norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$. Observe that
$\frac{\ip{v_i}{x}}{\gamma} \in [-\frac{1}{\gamma}, \frac{1}{\gamma}]$. Consider
$S_{s,r}\left(\frac{\ip{v_i}{x}}{\gamma}\right)$ for any $i \in
\{1,2,\dots,m\}$. Parts 1,2, and 3 of \autoref{lemma:properties-of-s-n-r}
and the fact $1/\gamma \le 2^s$ imply that
$S_{s,r}\left(\frac{\ip{v_i}{x}}{\gamma}\right) \le 1 +
\frac{1}{2m}$ for all $i \in \{1,2,\dots,m\}$. By the choice of $x$, there
exists $j \in \{1,2,\dots,m\}$ such that $\ip{v_j}{x} \le - \gamma$. Part 2 of
\autoref{lemma:properties-of-s-n-r} implies that
$S_{s,r}\left(\frac{\ip{v_j}{x}}{\gamma}\right) \in [-1-\frac{1}{2m},-1]$. Thus,
\begin{align*}
Q(x)
& = \left( \sum_{i=1}^m S_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right) \right) - \left(m - \frac{1}{2} \right) \\
& = S_{s,r}\left( \frac{\ip{v_j}{x}}{\gamma} \right) + \left( \sum_{\substack{i ~:~ 1 \le i \le m \\ i \neq j}} S_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right) \right) - \left(m - \frac{1}{2} \right) \\
& \le -1 + (m-1) \left( 1 + \frac{1}{2m} \right) - \left(m - \frac{1}{2} \right) \\
& \le 1/2 \; .
\end{align*}
\end{proof}

To prove parts 1 and 2 of \autoref{theorem:polynomial-approximation-1}
first note that part 4 of \autoref{lemma:properties-of-s-n-r} implies that
for any $x$ such that $\norm{x} \le 1$,
$B_{s,r}\left( \frac{\ip{v_i}{x}}{\gamma} \right)$ is positive.
Thus $p(x)$ and $Q(x)$ have the same sign on the unit ball.
\autoref{lemma:properties-of-q} states that $Q(x) \ge 1/2$
and $Q(x) \le -1/2$ in the two respective regions of the unit ball.
It remains to show that $|p(x)| \ge m^{s^2m/2}$ in those regions.
Consider any $x$ in either
$\displaystyle \bigcap_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \ge \gamma \right\}$
or in $\displaystyle \bigcup_{i=1}^m \left\{ x \in \R^d ~:~ \norm{x} \le 1, \ \ip{v_i}{x} \le - \gamma \right\}$.
Then,
\begin{align*}
|p(x)|
& = |Q(x)| \prod_{j=1}^m B_{s,r} \left( \frac{\ip{v_j}{x}}{\gamma} \right) \\
& \ge |Q(x)| \left( 2^{s(s+1)r} \left( 1 - \frac{1}{4m+1} \right) \right)^m \\
& \ge |Q(x)| \cdot 2^{s(s+1)rm} e^{-4} && \text{(since $\left(1-\frac{1}{4m+1}\right)^m \ge e^{-4}$)} \\
& \ge 4 |Q(x)| \cdot 2^{s^2rm} && \text{(since $e^{-4} \ge 1/2$ and $2^{rms} \ge 8$)} \\
& \ge 4 |Q(x)| \cdot (4m+1)^{s^2m/2} && \text{(since $r \ge 1 + \frac{1}{2} \log_2(4m+1)$)} \\
& \ge 2 |Q(x)| \cdot m^{s^2m/2} \\
& \ge m^{s^2m/2} && \text{(\autoref{lemma:properties-of-q})} \; .
\end{align*}
TODO
